Open Science, Open Plant Pathology

### 2020-04-21 analysis/paper/paper.Rmd
Flesh out methods and results from article surveys

@@

Twenty-one plant pathology discipline journals were selected by the authors as representations of discipline-based journals target by the plant pathology research community.
Among them, both fundamental and/or applied as well as journals covering specific group of pathogens/plants or broad areas were included. Two hundred articles were randomly selected from issues published from 2012 to 2018.
A list of randomly selected pages was assigned to a randomized list of the 21 journals [@Sparks2017] where the page number fell within an article for the given journal.
In cases where an article was not suitable, _e.g._, a review or otherwise not related to plant pathology, the next article was selected until a suitable article was found. Notes regarding the selection of articles can be found in the file, XXXX, available in this paper's repository.
The pages list was numbered from page one and went to 150.
This was done since some journals restart their numbering with each issue and also ensures that the journal is more likely to have a page number corresponding to the randomly generated value.
This also assumes that there is no effect or bias on reproducibility based on the time of year that an article was published, since most journals start with page number one at the beginning of the year.
The list of journals was saved as a comma separated value (CSV) file and imported into R [@RCT2020].

@@

##### pagebreak

>>>>

## Article Selection

In order for us to determine where we are as a discipline with regards to openly and reproducibly sharing data, we randomly selected 300 articles published from 2012 to 2018 across 21 journals (Table \@ref(tab:journals)) representative of the pathology discipline, and assigned them scores reflecting their openness and reproducibility.
Among the journals, both fundamental and/or applied as well as journals covering specific group of pathogens/plants or broad areas were included.
Three hundred articles were randomly selected from issues published from 2012 to 2018.
A list of randomly selected pages was assigned to a randomized list of the 21 journals [@Sparks2017] where the page number fell within an article for the given journal.
In cases where an article was not suitable, _e.g._, a review or otherwise not related to plant pathology, the next article was selected until a suitable article was found.
Notes regarding the selection of articles can be found in the file, XXXX, available in this paper's repository.
The pages list was numbered from page one and went to 150.
This was done since some journals restart their numbering with each issue and also ensures that the journal is more likely to have a page number corresponding to the randomly generated value.
This also assumes that there is no effect or bias on reproducibility based on the time of year that an article was published, since most journals start with page number one at the beginning of the year.
The list of journals was saved from Google's gsheets as a comma separated value (CSV) file for later use and distribution.

## Scoring criteria

Four individuals were assigned to rate a randomised list of journal articles.

Journals were provided a score for whether they were completely open (TRUE), behind a paywall (FALSE) or a combination (BOTH).
The five-year impact factor for each journal was retreived from (XXXXX) and entered in the notes.

Articles were provided classifications, whether they were fundamental or applied and a note was made whether the article's primary focus was molecular (TRUE/FALSE).

Where possible the software used was recorded in the notes if it was cited or otherwise specified in the article text.

Articles were scored in four areas for reproducibility on a scale of 0 to 3:

  * **computational methods available**

    * 0 - Not available or not mentioned in the publication

    * 1 - Uses expensive proprietary software that only institutions would typically purchase (_i.e._, 
    ArcGIS standard is $7000)

    * 2 - Uses proprietary software that most individuals can afford (_e.g._, Excel, Matlab? ($500))

    * 3 - Uses entirely open source and free software (_e.g._, R, Python)

    * NA - No software was used in the research that can be determined as the article is written

  * **software available**

    * 0 - Not available or not mentioned in the publication

    * 1 - Uses expensive proprietary software that only institutions would typically purchase (_i.e._, 
    ArcGIS standard is $7000)

    * 2 - Uses proprietary software that most individuals can afford (_e.g._, Excel, Matlab? ($500))

    * 3 - Uses entirely open source and free software (_e.g._, R, Python).

    * NA - No software was used in the research that can be determined as the article is written

  * **software cited**

    * 0 - not mentioned

    * 1 - Software mentioned by name only

    * 2 - Software cited with version number

    * 3 - All software components (SAS PROCs, R or Python packages) etc. cited

    * NA - No software was used in the research that can be determined as the article is written.

  * **data available**

    * 0 - Not available or not mentioned in the publication

    * 1 - Available upon request to author

    * 2 - Online, but inconvenient or non-permanent (_e.g._, login needed, pay wall, FTP server, personal 
    lab website that may disappear, or may have already disappeared)

    * 3 - Freely available online to anonymous users for foreseeable future (_e.g._, archived using Zenodo, 
    dataverse or university library or some other proper archiving system)

## Results

No journals were completely closed, all offered at least an option for open access but some were completely open.

The majority of the articles were classed as fundamental, 160, with 138 classed as applied.

Most articles did not make any computational methods available in any fashion with two classing as "3" (0.007&nbsp;%), which was the highest score available, 287 were classed as "0" (96&nbsp;%) and 9 articles (0.03&nbsp;%) appeared to not use any computational methods.

More articles did a good job of using software that was reasonably available to anyone with 206 scoring "2" or "3" (Table \@ref(tab:sa)).

A custom R function, `Reproducibility.in.Plant.Pathology::import_rrpp()`, was written to import the data, format the columns properly and calculate the overall reproducibility score in R [@RCT2020].

The majority of articles scored very low in our reproducibility assessment, 82.6&nbsp;% scored 0 out of 12, which was not reproducible at all (Table \@ref(tab:reproducibility)).
The highest scoring article recieved an 11 out of 12.

@@

@@

# Tables

```{r journals, echo=FALSE, message=FALSE, warning=FALSE}
library("dplyr")
library("janitor")
library("flextable")
library("officer")

import_rrpp() %>%
  tabyl(journal) %>%
  select(-percent) %>%
  flextable() %>%
  set_caption(
    "Table: (\\#tab:journals) Journal titles selected for inclusion and the number (n) of articles from each journal that were evaluated."
  ) %>% 
  autofit()
```

```{r software_avail, echo=FALSE, message=FALSE, warning=FALSE}
import_rrpp() %>%
  tabyl(software_avail) %>%
  select(-percent) %>% 
  rename(`software availability` = software_avail,
         percent = valid_percent) %>% 
  flextable() %>%
  set_caption(
    "Table: (\\#tab:sa) Article ratings for software availability where '0', not available or not mentioned in publication to '3', preferably open source and freely available."
  ) %>% 
  autofit()
```

```{r reproducibility_score, echo=FALSE, message=FALSE, warning=FALSE}
import_rrpp() %>% 
  tabyl(reproducibility_score) %>% 
  adorn_pct_formatting() %>% 
  rename(score = reproducibility_score) %>% 
  flextable() %>% 
  set_caption("Table: (\\#tab:reproducibility) Tally of reproducibility scores where 0 is not reproducible and 12 is fully reproducible, _e.g._ all software cited, software used are freely available to anyone, computational methods (statistcal analysis or other computational code) are freely available for download and use, and raw data are available for download and use.") %>% 
    autofit()

```
